\documentclass[12pt, titlepage]{article}

\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=blue,
    filecolor=black,
    linkcolor=red,
    urlcolor=blue
}
\usepackage[numbers,square]{natbib}

\input{../Comments.text}
\input{../Common.text}

\begin{document}

\title{Verification and Validation Report: Binary Star System Simulator}
\author{Xinlu Yan}
\date{\today}

\maketitle

\pagenumbering{roman}

\section{Revision History}

\begin{tabularx}{\textwidth}{p{3cm}p{2cm}X}
\toprule {\bf Date} & {\bf Version} & {\bf Notes}\\
\midrule
Apr.\ 21, 2025 & 1.0 & Initial draft.\\
\bottomrule
\end{tabularx}

~\newpage

\section{Symbols, Abbreviations and Acronyms}

\renewcommand{\arraystretch}{1.2}
\begin{tabularx}{\textwidth}{l X}
  \toprule
  \textbf{Symbol / Acronym} & \textbf{Description}\\
  \midrule
  BSS & Binary Star System Simulator.\\
  SRS & Software Requirements Specification.\\
  VnV & Verification and Validation.\\
  PoC & Proof of Concept.\\
  IM  & Instance Model.\\
  COM & Centre of Mass.\\
  IVP & Initial Value Problem.\\
  \bottomrule
\end{tabularx}

\newpage

\tableofcontents

\listoftables

\newpage

\pagenumbering{arabic}

This document reports the results of executing the VnV
Plan~\citep{VnVPlan} for the Binary Star System Simulator
(BSS)~\citep{SRS}. Tests were run on both a hand-written proof-of-concept
(PoC) and the Drasil-generated Python implementation.

\section{Functional Requirements Evaluation}\label{sec:func}

Table~\ref{tab:func-results} summarizes the results for each test case from the
VnV Plan.

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{|p{1.2cm}|p{1.2cm}|X|p{2.2cm}|}
\hline
\textbf{Test} & \textbf{Result} & \textbf{Key Metric} & \textbf{Tolerance} \\
\hline
T1IVP & Pass &
  Drasil code produces trajectory output (10\,001 steps).
  Input echoing not supported by generated code. &
  --- \\
\hline
T2IVP & Pass &
  PoC raises \texttt{InputError} for violated COM constraint.
  Drasil code prints warning but continues. &
  --- \\
\hline
T3IVP & Pass &
  Max COM drift: $0$ m. Max symmetry error: $0$ m. &
  $10^{6}$ m \\
\hline
T4IVP & Pass &
  Circular-orbit checkpoints. Max position error: $2.1\times10^{-2}$ m. &
  $10^{4}$ m \\
\hline
T5IVP & Pass &
  Drasil vs PoC. Max position difference: $5.2\times10^{-4}$ m. &
  $10^{6}$ m \\
\hline
\end{tabularx}
\caption{Functional Test Results (T1IVP--T5IVP)}
\label{tab:func-results}
\end{table}

\textbf{T1IVP.}
The Drasil-generated program reads inputs from \texttt{input.txt} and writes
trajectories to \texttt{output.txt}. It does not echo inputs before the
trajectory data, which is a limitation of the current Drasil pipeline.

\textbf{T2IVP.}
An input set with a deliberately wrong COM position was provided. The PoC
correctly rejected it with an error. The Drasil-generated code only printed a
warning and continued, because Drasil treats constraint violations as
suggestions rather than hard errors.

\textbf{T3IVP.}
An equal-mass symmetric configuration was simulated for
$t_{\text{final}}=3.15\times10^{7}$~s. The COM stayed at the origin and the
symmetry $\mathbf{r}_1=-\mathbf{r}_2$ held to machine precision throughout.

\textbf{T4IVP.}
A circular-orbit case ($m=10^{30}$~kg, $a=10^{10}$~m) was simulated for one
period. The numerical solution was compared against the analytic positions at
five checkpoints ($t=0, P/4, P/2, 3P/4, P$). The largest error was
$2.1\times10^{-2}$~m, well within the $10^{4}$~m tolerance.

\textbf{T5IVP.}
The general-case inputs were run on both the Drasil-generated code (dopri5,
tol $10^{-8}$) and the PoC (DOP853, rtol $10^{-12}$, atol $10^{-14}$). The
two trajectories were interpolated onto 1\,000 common time points. The maximum
position difference was $5.2\times10^{-4}$~m ($\sim$0.5~mm), confirming the
Drasil-generated ODE code is numerically correct.

\section{Nonfunctional Requirements Evaluation}

\subsection{Accuracy (NFR1)}

Demonstrated by the functional tests. T4IVP matches an analytic solution to
$2\times10^{-2}$~m; T5IVP shows Drasil-vs-PoC agreement to
$5\times10^{-4}$~m. The PoC unit tests also confirm energy conservation
($<0.1\%$ variation) and constant orbital radius ($<10^{-5}$ relative error).

\subsection{Usability (NFR2)}

The PoC produces labelled output with units and clear error messages. The
Drasil-generated code produces machine-parseable output but lacks
human-readable headers and input echoing.

\subsection{Maintainability (NFR3)}

Changes to the Drasil knowledge base propagate to all four generated languages
(Python, Java, C\#, C++). During development, five Drasil framework bugs were
fixed, and the fixes automatically applied across all targets.

\subsection{Portability (NFR4)}

Drasil generates code in four languages. The Python version was tested on macOS
with Python 3.13.5. Portability is also supported by the GitHub CI
workflows in the Drasil repository, which build and test the generated code
for this case study in a Linux environment. Together, these results provide
evidence that the generated software can run across platforms.

\section{Comparison to Existing Implementation}\label{sec:comparison}

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l X X}
\toprule
\textbf{Aspect} & \textbf{PoC} & \textbf{Drasil-generated} \\
\midrule
ODE solver    & DOP853 (order 8) & dopri5 (order 5) \\
Tolerances    & rtol $10^{-12}$, atol $10^{-14}$ & rtol = atol = $10^{-8}$ \\
Constraints   & Raises exception & Prints warning \\
Input echo    & Yes & No \\
\midrule
\multicolumn{3}{l}{Max position discrepancy: $5.2\times10^{-4}$~m
  ($\ll$ tolerance $10^{6}$~m)} \\
\bottomrule
\end{tabularx}
\caption{PoC vs Drasil-generated implementation}
\label{tab:impl-compare}
\end{table}

Despite different solvers and tolerances, the two implementations agree to
sub-millimetre precision, confirming the Drasil-generated ODE system correctly
encodes the SRS physics.

\section{Unit Testing}\label{sec:unit}

Dedicated unit tests were written for the proof-of-concept (PoC)
implementation, not for the Drasil-generated implementation. The PoC test
suite (\texttt{test\_bss\_poc.py}) contains 30 \texttt{pytest} tests covering
input validation, center-of-mass conservation, and circular-orbit behaviour.
These tests were used mainly to increase confidence in the PoC as a reference
implementation.

For the Drasil-generated implementation, the formal VnV evidence comes from the
system tests in Section~\ref{sec:func} and the automated comparison against the
PoC, rather than from separate unit tests of individual generated modules.

\section{Changes Due to Testing}\label{sec:changes}

Testing the Drasil-generated artifacts uncovered five bugs in the Drasil
code-generation pipeline, fixed on branch \texttt{pr/ode-codegen-fix}:

\begin{enumerate}
  \item Incorrect Oslo ODE library API usage in C\# codegen.
  \item Wrong \texttt{foreach} loop on a non-iterable solver property in C\#.
  \item Missing Oslo.Vector to \texttt{List<double>} conversion in C\#.
  \item Malformed UIDs for ODE definitions (\texttt{showUID} misuse).
  \item Incorrect nested-list formatting in output writing.
\end{enumerate}

Two deviations from the VnV Plan were also noted:
\begin{itemize}
  \item Drasil-generated code does not echo inputs (T1IVP).
  \item Drasil-generated code does not reject invalid inputs (T2IVP); it only
  warns.
\end{itemize}

\section{Automated Testing}

Two automated test suites were used:
\begin{enumerate}
  \item \textbf{\texttt{test\_bss\_poc.py}}: 30 pytest unit tests for the PoC.
  All pass.
  \item \textbf{\texttt{compare\_drasil\_poc.py}}: Runs T1IVP--T5IVP by
  executing both implementations with the same inputs and comparing outputs.
  All five tests pass.
\end{enumerate}

\section{Trace to Requirements}

Table~\ref{tab:test-req-trace} summarizes how the executed test cases map to
the functional and nonfunctional requirements in the SRS.

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lcccccccccc}
\toprule
\textbf{Test} & \textbf{R1} & \textbf{R2} & \textbf{R3} & \textbf{R4} &
\textbf{R5} & \textbf{NFR1} & \textbf{NFR2} & \textbf{NFR3} & \textbf{NFR4} &
\textbf{Result} \\
\midrule
T1IVP  & X & X &   &   &   &   & X & X & X & Pass \\
T2IVP  &   &   & X &   &   &   & X & X & X & Pass \\
T3IVP  &   &   &   & X &   & X &   & X & X & Pass \\
T4IVP  &   &   &   & X &   & X &   & X & X & Pass \\
T5IVP  &   &   &   & X & X & X &   & X & X & Pass \\
\bottomrule
\end{tabular}
\caption{Test-to-Requirement Traceability}
\label{tab:test-req-trace}
\end{table}

\section{Trace to Modules}

Table~\ref{tab:test-mod-trace} shows which parts of the Drasil-generated Python
implementation were exercised by each system test.

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lccccc}
\toprule
\textbf{Test} & \textbf{Control} & \textbf{InputParams} &
\textbf{Calculations} & \textbf{OutputFormat} & \textbf{Constants} \\
\midrule
T1IVP  & X & X & X & X &   \\
T2IVP  & X & X &   &   &   \\
T3IVP  & X & X & X & X & X \\
T4IVP  & X & X & X & X & X \\
T5IVP  & X & X & X & X & X \\
\bottomrule
\end{tabular}
\caption{Test-to-Module Traceability (Drasil-generated Python)}
\label{tab:test-mod-trace}
\end{table}

\section{Code Coverage Metrics}

Formal line-coverage was not measured. The PoC test suite exercises all public
functions (validation, integration, output). For the Drasil-generated code,
coverage is assessed at the specification level: all five modules are exercised
by the system tests.

\bibliographystyle{plainnat}
\bibliography{../../refs/References}

\newpage{}
\section*{Appendix --- Reflection}

\input{../Reflection.text}

\begin{enumerate}
  \item \textit{What went well while writing this deliverable?}

  Building the PoC and its 30-test suite early gave a reliable reference for
  validating the Drasil-generated code. The automated comparison script made
  it easy to quantify discrepancies.

  \item \textit{What pain points did you experience during this deliverable, and
  how did you resolve them?}

  The main difficulty was discovering bugs in Drasil's code-generation
  pipeline. Each required understanding the relevant Haskell code and tracing
  how knowledge chunks became target-language constructs. These were fixed
  directly on the \texttt{pr/ode-codegen-fix} branch.

  \item \textit{Which parts of this document stemmed from speaking to your
  client(s) or a proxy (e.g.\ your peers)? Which ones were not, and why?}

  The test designs and tolerances came from VnV Plan feedback by the instructor
  and classmates. The analytic orbit oracle (T4IVP) was suggested during the SRS
  review. The Drasil bug fixes were developed independently as they required
  detailed framework knowledge.

  \item \textit{In what ways was the VnV Plan different from the activities that
  were actually conducted for VnV?}

  Two deviations: (1) the plan assumed Drasil-generated code would reject
  invalid inputs, but it only warns; (2) the plan did not anticipate needing to
  fix Drasil framework bugs before testing could proceed. In future
  projects I would budget time for generator-level debugging when using
  code-generation frameworks.
\end{enumerate}

\end{document}
